### 2020-12-16 ch/1-introduction.tex
Meager progress

Now, one possible objection here is:
if those datasets are so bad,
then maybe we just should not train any \ac{ML} model on these and should not use them to make automated decisions.
While this question is mostly beyond the scope of this document, let me offer some thoughts on this:
It's true that even after the application of de-biasing techniques,
the resulting models still should not be fully trusted,
but they can still ease the burden of checking everything manually;
similar to an email spam filter which is not perfect, but still very useful.
One could imagine a hybrid approach where an automated system makes preliminary decisions,
but random samples are reviewed by humans and decisions can always be challenged.
Ideally, the model itself would tell us about decisions it is uncertain about.

@@

For \emph{sampling bias},
there is also a special attribute $s$, but it does not necessarily have to denote a \emph{sensitive} attribute;
it can more generally be a \emph{spurious} variable which is correlated with the class label $y$ in the training set,
but is not truly predictive of $y$ in the general case.
Sampling bias then means that the training set is not uniformly sampled from the underlying distribution,
instead, sampling depends on $s$ and $y$;
\eg, there might be almost no samples of $s=0$ and $y=1$ in the training set.

\thomas{should there be something about fairness metrics here?}

\thomas{Possible segway:}
While we could determine how strongly a classifier has been affected by these biases
by computing their accuracy on an unbiased test set,
there are more

>>>>

Now, one possible objection here is:
if those datasets are so bad,
then maybe we just should not train any \ac{ML} model on these and should not use them to make automated decisions.
While this question is mostly beyond the scope of this document, let me offer some thoughts on this:
It's true that even after the application of de-biasing techniques,
the resulting models still should not be fully trusted,
but they can still ease the burden of checking everything manually;
similar to an email spam filter which is not perfect, but still very useful.
Or, put another way, it is always important to check what the realistic alternative is;
we should not compare a model to a non-existing perfect ideal, but to the actual thing that would be used instead.
One could imagine a hybrid approach where an automated system makes preliminary decisions,
but random samples are reviewed by humans and decisions can always be challenged.
Ideally, the model itself would tell us about decisions it is uncertain about.

@@

For \emph{sampling bias},
there is also a special attribute $s$, but it does not necessarily have to denote a \emph{sensitive} attribute;
it can more generally be a \emph{spurious} variable which is correlated with the class label $y$ in the training set,
but is not truly predictive of $y$ in the general case.
Sampling bias then means that the training set is not uniformly sampled from the underlying distribution,
instead, sampling depends on $s$ and $y$;
\eg, there might be almost no samples of $s=0$ and $y=1$ in the training set.
The effect is that \ac{ML} models are tempted to take shortcuts and use $s$ as a shorthand for $y$.

As the training set is biased in these settings,
evaluation metrics like accuracy should be computed with respect to an \emph{unbiased} test set.
However, such a set is often not available or has to be constructed, which leaves open the question how to construct it.
Furthermore, the overall accuracy often does not capture all that we care about.
For these and other reasons,
it is usually necessary to go beyond metrics like accuracy, when evaluating models in these settings.
Concretely, we use \emph{fairness metrics} which are based on certain definitions of fairness.
These fairness metrics can be computed for predictions on a \emph{biased} test set
and still give us an idea of the bias in the model.
Even with an \emph{unbiased} test set, these metrics are useful,
because they tell us in what specific way, if any, a model has failed to correct for the training set bias.

The simplest fairness definition is \ac{DP}, also called statistical parity or independence.
